---
title: "Project Code"
output: pdf_document
author: "Jack Nowacek"
date: "2023-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages

```{r packages}
library(tidyverse)
```

## Data Introduction

```{r Data Formatting}
data <- Shorebird_Data |>
  slice(1:23)
```

Is shorebird nest initiation driven by climatic conditions, food or both?

  linear regressions between food emergence day/abundance and bird nesting day
  same with spring temp/snow cover
  side-by-side line and point graphs of emergence days over the years, do they respond each year?        respond to previous years?
  
Are there differences among shorebird species?

  simpler stats, just a t-test of the average or median emergence, are they statistically different?
  do they respond to each of the previous tests to a greater/lesser degree

Are there indications of a temporal mis-match between shorebirds and their prey?

  answered from the previous plots
  
Test 1: linear regression between the emergence days, maybe a plot with ab lines

Test 2: side-by-side point graphs, look at trends visually to see if there is clear following or response

Test 3: check for stat sig diff btw each species, see what they respond to, hypothesize

## EDA

```{r exploratory data analysis}
data |>
  ggplot(aes(x = Year, group = 1, color = SpringTemp)) +
  geom_path(aes(y = SpringSnow)) +
  geom_point(aes(y = SpringSnow)) +
  theme_linedraw() +
  labs(title = "Spring Snow Coverage", y = "Snow Cover Percentage")
```

```{r day of year plots}
# data |>
#   ggplot(aes(x = Year, group = 1)) +
#   geom_path(aes(y = c()) +
#   geom_point(aes(y = SpringSnow)) +
```

EOF